---
title: "CIEE Productivity and Reproducibility Draft Manuscript"
author: "Hannah Marton"
date: "2025-09-11"
output: pdf_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(tinytex)
library(lterdatasampler)
```

# Introduction

This is my draft manuscript for the my CIEE Productivity and Reproducibility project. I am learning how to write text in "Markdown" format. I can write in **bold** and *italics*.

Acid rain leaches calcium out of the soil, a nutrient critical to the growth of the sugar maple (*Acer saccharum*). This lack of calcium leads to the soil turning more acidic, and creates a stressful environment for the species [@ramanujan2006]. We will use the Hubbard Brook Experimental Forest Sugar Maples ("hbr_maples") dataset from the 'lterdatasampler' package to explore the addition of calcium on sugar maple seedling growth.

# Methods

## Dataset

This project uses data from 'lterdatasampler', an R package that provides sampler datasets for teaching and learning purposes [@horst2023].

The package contains several sample datasets, but we will work with the "hbr_maples" data. "[This] dataset contains observations on sugar maple seedlings in untreated and calcium-amended watersheds at Hubbard Brook Experimental Forest in New Hampshire" [@horst2023].

The data originates from "Health and mycorrhizal colonization response of *Acer saccharum* seedlings to calcium addition in Watershed 1 at the Hubbard Brook Experimental Forest" [@juice2019].

```{r readrawdata, include=FALSE}
data("hbr_maples")
write.csv(hbr_maples, "../00_rawdata/hbr_maples.csv",
          row.names = FALSE)
```

## Analysis

I can do a small data transformation for practice here.

\newpage

# Results

Let's make a simple plot that compares sugar maple seedling height in the untreated and calcium treated watersheds.

```{r ggplot, echo=FALSE, fig.cap="*Acer saccharum* seedling height in untreated (Reference) and treated (Calcium addition) watersheds.", fig.pos='H'}
library(tidyverse)
ggplot(hbr_maples, aes(x = watershed, y = stem_length, colour = watershed)) +
  geom_jitter(
    aes(color = watershed),
    alpha = 0.5,
    show.legend = FALSE,
    position = position_jitter(width = 0.2, seed = 0)) +  
  geom_boxplot(alpha = 0.2, outlier.shape = NA) +     
  facet_wrap(~ year) +
  theme_bw() +
  labs(x = "Watershed", y = "Stem length (mm)") +
  scale_x_discrete(labels = c("W1" = "Calcium addition", "Reference" = "Reference")) +
  scale_color_manual(values = c("W1" = "springgreen4", "Reference" = "peachpuff4")) 
```

# Discussion

This is where I'd write my discussion. I can add more citations here [@peters2004; @juice2006].

\newpage
# References
